---
title: "Predictive Modeling of Annual Inflation Using Weighted Least Squares Regression"
author: "Annisa Indriani"
date: "2024-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Library

```{r}
library(readxl)
library(lmtest)
library(nortest)
library(car)
library(dplyr)
library(ggplot2)
library(reshape2)
library(performance)
library(gamlr)
```

## Import Data

```{r}
gdp = read_excel("Dataset Kesejahteraan Sosial dan Ekonomi Negara Asia 2022.xlsx",
                 sheet = 2)
gdp
```

## Exploratory Data Analysis

### Check Data Structure

```{r}
str(gdp)
```

### **Check Summary**

```{r}
summary(gdp)
```

### **Check Missing Values**

```{r}
colSums(is.na(gdp))
```

### **Correlation Heatmap**

```{r}
correlation_matrix <- cor(gdp %>% select_if(is.numeric))

melted_correlation <- melt(correlation_matrix)
ggplot(data = melted_correlation, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = 'blue', high = 'red', mid = 'white', 
                       midpoint = 0, limit = c(-1, 1), name = 'Correlation') +
  theme_minimal() +
  labs(title = 'Correlation Heatmap')
```

## **Linear Regression Model**

```{r}
model1 = lm(Inflation_Consumer_Prices_Annual ~ 
              Households_Consumption_Expenditure_Annual_Growth +
              Age_Dependency_Ratio +
              Inflation_GDP, data = gdp)
summary(model1)
```

From the summary above, we can see that the model is significant because the p-value is less than alpha 0.05. In addition, the variables also have p-values below 0.05, indicating a relationship between Households Consumption Expenditure Annual Growth, Age Dependency Ratio, Inflation GDP, and Inflation Consumer Prices Annual.

## **Assumptions Test**

### **Independent**

```{r}
dwtest(model1)
```

The Durbin-Watson test shows that the model has a p-value of more than alpha 0.05, which indicates that the independence assumption is met.

### **Identic**

```{r}
bptest(model1)
```

The Breusch Pagan test shows that the model has a p-value of less than 0.05, which indicates that the identical assumption is not met, and therefore, the model cannot be used for further analysis. Thus, a weighted least squares regression model will be used to deal with this issue.

### **Normality**

```{r}
e1 <- residuals(model1)
shapiro.test(e1)
```

The Shapiro-Wilk test shows that the model has a p-value of more than alpha 0.05, which indicates that the normality assumption is met.

### **Multicollinearity**

```{r}
vif(model1)
```

A VIF value of less than 10 indicates that there is no multicollinearity between Households Consumption Expenditure, Annual Growth, and Age Dependency Ratio.

## **Weighted Least Squares Regression Model**

```{r}
weight1 = 1 / lm(abs(model1$residuals) ~ model1$fitted.values)$fitted.values^2

model_wls1 <- lm(Inflation_Consumer_Prices_Annual ~ 
                   Households_Consumption_Expenditure_Annual_Growth +
                   Age_Dependency_Ratio + Inflation_GDP, data = gdp, 
                 weights = weight1)
summary(model_wls1)
```

From the summary above, we can see that the model is significant as the p-value is smaller than alpha 0.05. In addition, all variables except Households Consumption Expenditure Annual Growth have a p-value below 0.05, which means there is a relationship between ram, battery, rear cam and phone price.

```{r}
model2 = lm(Inflation_Consumer_Prices_Annual ~
              Age_Dependency_Ratio +
              Inflation_GDP, data = gdp)

weight2 = 1 / lm(abs(model2$residuals) ~ model2$fitted.values)$fitted.values^2

model_wls2 <- lm(Inflation_Consumer_Prices_Annual ~ 
                   Age_Dependency_Ratio + Inflation_GDP, data = gdp, 
                 weights = weight2)
summary(model_wls2)
```

From the summary above, we can see that the model is significant because the p-value is less than alpha 0.05. In addition, the variables also has a p-value below 0.05, which means there is a relationship between Age Dependency Ratio, Inflation GDP, and Inflation Consumer Prices Annual.

### Breusch-Pagan Test

```{r}
bptest(model_wls2)
```

The Breusch-Pagan test shows that the model has a p-value of more than alpha 0.05, which indicates that the independence assumption is met.

## **Goodness of the Model**

### **AIC**

```{r}
AIC(model1)
AIC(model_wls1)
AIC(model_wls2)
```

### RSE

```{r}
summary(model1)$sigma
summary(model_wls1)$sigma
summary(model_wls2)$sigma
```

Based on the AIC and RSE values, it can be seen that the WLS model with significant variables has a smaller value than other models. Thus, the model is the best model and can be used for analysis.
